---
title: "Ecology of the Thymus"
author: "Mary Browning, John Wares, Andrew Park, Nancy Manley"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: pdf_document
bibliography: thymus.bib
---

<<<<<<< HEAD
## Introduction2

Abstract

The correct organization and abundance of cell types is important for optimal organ performance.  The ability to identify recurrent or cryptic cellular organizational patterns could help us to better understand tissue function and will enable us to make comparisons with diseased tissues.  However, quantitative approaches for tissue organization do not currently exist. Here, in an attempt to overcome this barrier, we have developed  novel molecular and computational analyses that allow us to study the organ of our choice, the thymus, in a statistical framework. First, we identify cell type abundance and location and use this information to recreate a virtual thymus section.  We then treat cell types like species in an ecosystem, thus enabling us to directly apply quantitative theory used in ecology to our virtual section. We use K means clustering to computationally produce regions of the thymus a priori, followed by Bray Curtis Dissimilarity Index to determine the degree of similarity between thymus samples and regions of the thymus. The output generated from these analyses can then be used to better understand spatial patterns/biogeography within a healthy functional thymus and will enable us to make informative comparisons of thymic tissues under different developmental, mutant, or disease conditions.

=======
## Introduction
>>>>>>> origin/master

Within the field of organ biology, there is a distinct lack of quantitative methods for analyzing the organization, as opposed to the composition, of organs and tissues.  Healthy organ function requires more than the appropriate proportions of component parts; these parts must also be arranged in specific ways. Without quantitative methods for assessing organization, a significant aspect of organ and tissue biology is inaccessible to analysis. However, beyond essentially descriptive assays, we have few theoretical tools available for interpreting organizational characteristics/patterns. In contrast, the fields of ecology and population genetics have a robust theoretical tradition, with multiple quantitative tools available for describing the natural world. Fortunately, a striking comparison can be drawn between the interactions that occur within organ systems and those that occur within ecosystems.  For instance, cell types in an organ, like species in an ecosystem, are influenced by the availability of space and resources as well as by the presence or absence of other cell types. The presence or absence of one cell type may alter the functional role and dominance of another, or the overall service the community provides to the larger system [e.g., @wootton05].  By comparing cell types within a biological organ to species in an ecosystem, quantitative methods used by ecologists can be directly applied to biological systems. Our overall goal is the development of a cellular ecology, where we can model cellular relationships and patterns in an effort to better understand the "biogeography" of the organ we wish to study. We aim to develop quantitative tools, modeled after the ecological toolkit, to spatially and statistically characterize healthy biological tissues, which is something that has never been done.  The results from this analysis will enable informative comparisons with tissues in states of disequilibrium, including disease or developmental states, in order to identify any recurring organizational differences.   

First, we aim to model a healthy adult organ. We will accomplish this by identifying cellular location and abundance.  We will then spatially and statistically characterize the organ by applying ecological theory directly to the produced model (something that has never been done before). By analyzing the exact spatial location and co-location of component cell types across the domain of a particular tissue or organ, we may gain a better understanding of which interactions - indicated by codistribution and sufficient density - can be validated by independent approach and analysis.  To an extent, we recapitulate the spatial scales described above: we explore the differentiation of distinct habitats within the environment (biogeography: identifying regions of endemicity or dramatically shifted abundance), the colocalization of particular sub-groups of cells (community ecology), and we use the tenets of macroecology to the extent that we can generalize that cells need a particular density to be viable interactors, and to the extent that common cells have a larger overall distribution (this is not tautological, this follows from the first: you cannot be rare and widespread in a functional sense).  The co-distribution of constituent species in a community is often used to indicate interaction - whether competitive, trophic, or facilitative [@verberk11;@angel11].

Specific structural aspects of organs with strong 3-D structures, such as the lung, kidney, and other organs that undergo branching morphogenesis, have been modeled with some success (2-4 below). However, the cellular organization of other types of organs can be particularly difficult to assess quantitatively.  For example, endodermal glandular organs are often organized at the cellular level, with organizational structures that are visible only with the use of molecular or cellular markers; thus detecting differences in organ structure depends in part on having distinct markers for known functional cell subsets.  Examples of this type of organ include the liver, pancreas, and thymus, all of which are of high clinical importance.  Developing a quantitative approach to assessing the structure and function of these organs is thus of direct biomedical relevance.  We performed our analysis on our model organ, the thymus.  The thymus was chosen because it is an excellent example of cellular level organization, with a strong connection between organization and function.  The thymus consists of developing T cells supported by a complex cellular environment containing a variety of resident cell types, including thymic epithelial cells (TECs), dendritic cells, vasculature, and mesenchymal cells.  These cell types comprise multiple microenvironments that direct and support thymocytes to develop from immature progenitors into mature T cells that are both self-tolerant and self-restricted.  T cell development in the thymus requires interactions with the thymic microenvironments that provide signals for their survival, proliferation, and differentiation [5].  Despite their critical role in the generation of cellular immunity and the clinical importance of thymic regeneration, the composition and organization of thymic microenvironments and the mechanisms that promote their proper development and function are not fully understood, in part due to a lack of technical and theoretical approaches for quantifying tissue-level properties.  Thus, the thymus has many characteristics that make it an excellent system for developing and testing quantitative modeling: a diverse cellular composition that can be identified with cell type-specific markers, regional organization that is required for maximal organ function,  genetic models with diverse effects on organ composition and function, assays for experimentally inducing organ degeneration and regeneration, and high biomedical relevance. To date, there are no quantitative theoretical models of thymus organ structure and function and no established methods for generating one.  We seek to develop a quantitative model of the organization of cell types that can be used to better understand thymic function as well as evaluate states of disequilibrium. The data obtained in this analysis will provide both a valuable resource for understanding the structure of a normal thymus, and a scientific framework and accessible tools for the analysis of mutant phenotypes in this project and in the field of organ biology. The new theoretical approaches that we will defive from these studies will also provide novel methods for assessing tissue organization in other organs, tissues, and disease states.

2. Hartmann D, Miura T (2007) Mathematical analysis of a free-boundary model for lung branching
morphogenesis. Mathematical medicine and biology : a journal of the IMA 24: 209-224.
3. Miura T (2008) Modeling lung branching morphogenesis. Current topics in developmental biology 81: 291-
310.
4. Oates AC, Morelli LG, Ares S (2012) Patterning embryos with oscillations: structure, function and dynamics
of the vertebrate segmentation clock. Development 139: 625-639.
5. Manley NR, Richie ER, Blackburn CC, Condie BG, Sage J (2012) Structure and function of the thymic
microenvironment. Front Biosci 17: 2461-2477.


Our approach is designed to take advantage of the cellular complexity of the thymus to provide a sensitive and quantitative method to measure changes in organ state. We use immunostaining on sagittal serial sections of a wildtype mouse thymus in order to identify distinct cellular subsets, followed by the use of novel computational approaches in order to quantitatively identify known and cryptic cellular spatial relationships. These methods are designed to replace the largely descriptive histological and IHC-based analysis currently available in the field, and to provide a more accurate and unbiased view of thymus biogeography.  First, we used the K means clustering algorithm to identify geographical clusters of cell types based on compositional similarities, position, and abundance. This approach has previously been successful in delineating distinct cell sub-population in tissue samples (6,7). K means clustering allows us to test the structure of thymus in a statistical framework that can provide unambiguous support for the number of compositionally distinct thymic regions which we can then compare to previously identified regions of the thymus. We then used a range of K values to compare to identify potentially cryptic organization. Our goal is not only to identify whether groups of 2 or more cell types have similar relative distributions (between regions or within the same regions), but also to quantify the extent to which mutants or diseased thymuses are significantly less structured than wild-type. This clustering algorithm generates a posterior probability that any subsample (quadrat) belongs to each of the k different groups, while maximizing the ratio of compositional differences between groups to compositional variation within groups.  The k groups produced are presumably functionally distinct, and a key hypothesis we will test is whether Kmutant is significantly less than kwildtype for each comparison. Identifying the aggregation characteristics of particular cell types or groups of cell types will allow us to quantify and validate results from the predictive models of thymus "cellular ecology" as described above.  

6. Veronika M, Evans J, Matsudaira P, Welsch R, Rajapakse J (2009) Sub-population analysis based on
temporal features of high content images. BMC Bioinformatics 10 Suppl 15: S4.
7. Cataldo S, Ficarra E, Macii E (2009) Automated Discrimination of Pathological Regions in Tissue Images:
Unsupervised Clustering vs. Supervised SVM Classification. In: Fred A, Filipe J, Gamboa H, editors.
Biomedical Engineering Systems and Technologies: Springer Berlin Heidelberg. pp. 344-356.)


In ecology, the identity and relative abundance of species can be compared across communities using various dissimilarity indicies (10. Magurran AE (2004) Measuring Biological Diversity. Malden, MA: Blackwell Publishing.) The Bray Curtis Dissimilarity Index in particular is a well characterized statistical method used to quantify the compositional dissimilarity between two sites, based on counts at each site [ Bray, J. R. and J. T. Curtis. 1957. An ordination of upland forest communities of southern Wisconsin. Ecological Monographs 27:325-349.]. We applied Bray Curtis to our samples to determine the degree of dissimilarity between clusters produced from K means clustering based on cell count. Bray Curtis was also used to make comparisons between different thymic samples and subsamples, and can also be compared to ecosystems/natural systems.

connected component labeling

We take this approach one step further by applying the same approaches described above to a mutant mouse thymus in order to determine if there are any detectable organizational differences compared to the wildtype. We selected the Aire knockout Mouse (Anderson MS; Venanzi ES; Chen Z; Berzins SP; Benoist C; Mathis D. 2005. The cellular mechanism of Aire control of T cell tolerance. Immunity 23(2):227-39. [PubMed: 16111640]  [MGI Ref ID J:100515]) for several reasons.  First, the Aire knockout is well characterized: DESCRIBE PHENOTYPE HERE?? there is debate as to whether mutant mice double in mTEC number (particularly with respect to K14) [26].  There is also debate as to whether dendritic cell organization and number is altered; there are differing reports in the literature [27,28]. From an ecological standpoint, we are interested in determining the degree to which Aire plays. Although we know Aire is functionally important, does it act as a Keystone species? Will the removal of Aire cause a shift BLAH. Or will there be no discernable organizational differences? 

26. Gillard GO, Dooley J, Erickson M, Peltonen L, Farr AG (2007) Aire-dependent alterations in medullary thymic epithelium indicate a role for Aire in thymic epithelial differentiation. J Immunol 178: 3007-3015.
27. Yano M, Kuroda N, Han H, Meguro-Horike M, Nishikawa Y, et al. (2008) Aire controls the differentiation program of thymic epithelial cells in the medulla for the establishment of self-tolerance. J Exp Med 205: 2827-2838.
28. Lei Y, Ripen AM, Ishimaru N, Ohigashi I, Nagasawa T, et al. (2011) Aire-dependent production of XCL1 mediates medullary accumulation of thymic dendritic cells and contributes to regulatory T cell development. J Exp Med 208: 383-394.


 
 Cell Types Used in Analysis
 SHOULD I INCLUDE PARAFFIN ANTIBODIES?
 ----------- ---------- -------------------------
  CELL TYPE    MARKER
  
  cTEC       CD205     

  mTEC       UEA1, Claudin3
             Claudin4, K14
             K5, Claudin5 
              
  Blood vessel Claudin5, CD31 
  
  Mesenchyme  PDGFRb
  
  Dendritic  CD11c
  
  Tcells     Foxp3, CD25  
                                       
----------- ---------- -------------------------
 
### Hypotheses

I predict that the distribution and abundance of cell types is sufficient for automated clustering of thymic regions.  The organization of a wild-type, mature thymus will either organize into more discrete regions than a mutant thymus, or the regionalization will be more variable and less discrete.
Next, we will examine cellular absence/presence with respect to other cell types to see if any patterns emerge
By using a well characterized dissimilarity index (Bray Curtis Dissimilarity Index), we will determine the degree of similarity between different regions and subregions.
We will apply the above approaches to a mutant thymus in order to see if orgazinational differences can be detected.

 
## Methods

### Terminology

I suggest an explanatory box for readers so that quadrat, section, composite, etc. can all be explained clearly in one place.

----------- ---------- -------------------------
  quadrat    spatial     sampled unit within
                          domain; cell types in
                          quadrats used to find
                          regions of similar
                          composition

  transect    spatial     linear array of 
                          sampled units 
                          (quadrats), may be
                          comprehensively or
                          randomly sampled
                          
  layer      tissue       actual tissue sample
                          used to assay 
                          single-cell thickness
                          composition
----------- ---------- -------------------------

### System and Data Collection

#### Mice
C57BL/6 male mice (N=4) at 5-6 weeks of age were purchased from Jackson Laboratories (Bar Harbor, Maine). Aire knockout (N=2) and WT male (N=2) mice at 5-6 weeks of age were obtained from Dr. Mark Anderson's lab at the University of California, San Francisco.

#### Tissue Preparation

Intact thymuses from the Anderson lab were removed and the right and left lobes were separated, placed in OCT compound, and immediately frozen before being shipped to the Manley lab. Mice were shipped from Jackson Laboratories and dissected and frozen in the Manley lab.

#### Antibodies

Primary antibodies used in this work include the biotinylated hamster CD11c (Cat#, 1:100), rabbit Claudin 3 (Cat#:  , 1:200), rabbit Claudin 4 (Cat#: , 1:200), supernatant hamster CD31 (Cat#, 1:50), rat CD25 (Cat#: , 1:50), biotinylated UEA1 (Cat#: , 1:200), conjugated mouse Claudin5 (Cat#: , 1:100), rabbit Keratin5 (Cat#, 1:200), biotin rat Foxp3 (Cat#, 1:50), goat Keratin14 (Cat#, 1:200), goat PDGFRb (Cat#, 1:100), and rat CD205 (Cat#, 1:200). 
Secondary antibodies were purchased from Invitrogen.
 
#### Frozen Sample Preparation
The entire left lobe was cut into 10-micron serial sagittal sections using a Leica CM3050 S cryostat. Sections were collected on glass slides and assigned a number corresponding to their location.  The middle third of these slides were selected for IHC.  Of these middle third, 9 sections from each sample that passed quality control standards were used for IHC.  These sections were fixed in -20 degree C acetone for 20 seconds immediately prior to application of blocking solution (10% donkey serum/PBS) for 30 minutes at room temperature.

#### IHC
Primary antibodies were mixed in PBS and incubated overnight at 4C.  The slides were rinsed with PBS 2x5 minutes and the secondary antibodies were mixed in PBS (1:800) and applied to the slides for 30 minutes.  Slides were rinsed with PBS 2x5 minutes, mounted in FluorGel (EMS) and coverslipped.   Marker combinations were determined by availability,  biological and technical considerations such as anticipated abundance, distribution of cell types, and reagent compatibility (antibody species of origin). 

#### Microscopy, PTGui, and CellProfiler
Multiple sections from each sample were used in the analysis (9 sections from each sample from the Anderson Lab; 8 sections from WT2, 6 from WT1).  The sections were assessed by specific quality control criteria such as lack of section flaws, clarity of image collection, and signal to noise ratio. Sections that passed the QC measures  were imaged at 20x as tiled, overlapping quadrats on the Zeiss  microscope and reconstructed using PTGui.  The image produced was optimized in Adobe Photoshop to reduce background and increase signal intensity so that cell counts could be taken as quickly and accurately as possible in CellProfiler.  The reconstructed image from PTGui was too large to work in CellProfiler, so the size of the image was decreased by taking the image width and dividing by 4. CellProfiler was chosen because it has the capability to quickly and accurately identify cells and output locations (x and y coordinates) in an excel file.  Results from the cell counts were  manually checked for accuracy and input into R for the spatial analysis.  Cell counts from serial sagittal sections stained with different marker combinations were collapsed into a single virtual section.  Our experimental appraoch is essentially 2-D.  Although the 3-D nature of the stromal network is an important aspect of its functionality [cite], a 2-D picture will provide an informative, if not complete view of the thymus.  After all, the qualitative assays currently used are all viewed in 2D sections and are the basis of our current views of thymus organization and function.   

####R
Files containing the X and Y coordinates from CellProfiler were uploaded into R and reformatted into a community matrix.  The K-means function in R's stats package was used to cluster the data. The  Bray-Curtis function in the Vegan package was used to calculate the Bray-Curtis index for the different regions produced from K-means clustering.  

####Aire
We will use an established mutant with specific effects on individual cell populations to test the effect on thymus organization. Null mutations in the Aire gene, which marks a specific subpopulation of mTECs, cause defective negative selection and multi-organ autoimmunity.  Analyses of TEC phenotypes have been performed, and there are reports that Aire mutants have defects in mTEC composition and organization (cite).  A recent report has also shown displaced dendritic cells in Aire mutant thymus (cite). These results have functional consequences for different models of Aire function as either a regulator of tissue-restricted antigens in the thymus or as a key modulator of mTEC differentiation and mechanistic implications for autoimmunity in these mice.  All of these reports point to the need for more robust quantitative methods to evaluate cellular organization.  The parallel in ecology is removal of a species from a community, which can lead to two distinct outcomes: the relative species abundance may not be affected if species are independently utilizing available resources, or the rand abundance of species may change dramatically, suggesting strong interactions depending on missind components [CITE]


Is there a way to keep multiple scripts for each sample and then call them in this script?
```{r use a shell type command to read in any *.txt, echo=FALSE, warning=FALSE, message=FALSE, results='hide',fig.show='hide'}

setwd("~/GitHub/Thymus/Datasets/Data with 11 cell types/Extra/Reprocessing_Mutant1/txt_2")

library(gplots)

msz=100  #matrix size in hist2d
cluster=4  #number of clusters produced for K means
txtfiles=list.files(pattern="*.txt")  #loads in order all files within folder

mat<-matrix(data=NA,nrow=10000,ncol=33) #this will change with size of matrix and # of sections/cell types

toFill<-as.data.frame(mat)  #changes to data frame

toUse0<-seq(1,33,1) #uses all 33 columns
toUse1<-seq(1,33,3) #uses columns that correspond to Section1
toUse2<-seq(2,33,3) #Section2
toUse3<-seq(3,33,3) #Section3

# define which vector to query here
#inLoop<-toUse1  #Determine what section you want to run here

for(i in toUse0){
  tmp = read.table(txtfiles[i], sep="\t", head=T)
  X=tmp[[1]]  #[,1]
  Y=tmp[[2]]
  my.xy<-hist2d(X,Y,nbins=c(msz,msz))
  obj <- my.xy$counts
  
  obj[1,1] = obj[1,1]-1
  obj[100,1] = obj[100,1]-1
  obj[1,100] = obj[1,100]-1
  obj[100,100]=obj[100,100]-1
  
  count<- as.vector(obj)
  toFill[,i]<-count
    
    }

for(i in toUse1){
  tmp = read.table(txtfiles[i], sep="\t", head=T)
  X=tmp[[1]]  #[,1]
  Y=tmp[[2]]
  my.xy<-hist2d(X,Y,nbins=c(msz,msz))
  obj <- my.xy$counts
  
  obj[1,1] = obj[1,1]-1
  obj[100,1] = obj[100,1]-1
  obj[1,100] = obj[1,100]-1
  obj[100,100]=obj[100,100]-1
  
  count<- as.vector(obj)
  toFill[,i]<-count
    
    }

for(i in toUse2){
  tmp = read.table(txtfiles[i], sep="\t", head=T)
  X=tmp[[1]]  #[,1]
  Y=tmp[[2]]
  my.xy<-hist2d(X,Y,nbins=c(msz,msz))
  obj <- my.xy$counts
  
  obj[1,1] = obj[1,1]-1
  obj[100,1] = obj[100,1]-1
  obj[1,100] = obj[1,100]-1
  obj[100,100]=obj[100,100]-1
  
  count<- as.vector(obj)
  toFill[,i]<-count
    
    }

for(i in toUse3){
  tmp = read.table(txtfiles[i], sep="\t", head=T)
  X=tmp[[1]]  #[,1]
  Y=tmp[[2]]
  my.xy<-hist2d(X,Y,nbins=c(msz,msz))
  obj <- my.xy$counts
  
  obj[1,1] = obj[1,1]-1
  obj[100,1] = obj[100,1]-1
  obj[1,100] = obj[1,100]-1
  obj[100,100]=obj[100,100]-1
  
  count<- as.vector(obj)
  toFill[,i]<-count
    
    }

result<-toFill[,toUse0]
result1<-toFill[,toUse1]
result2<-toFill[,toUse2]
result3<-toFill[,toUse3]

#print(results)

```

   
### Identification of Scale

As thymic epithelial cells vary in size from X to X', we considered the problem of how to generate sampled regions (quadrats) that contained sufficient information about the local community of cells that these regions could be classified. In other words, we need to have lots of data in each quadrat to have statistical power for clustering; we need lots of quadrats to improve the spatial resolution. Identifying the best scale is done how?

5. Quadrat Selection and Rarefaction: you need to explain how you did this, and the next chunk can include the code you used to achieve this. That way, our results can include one plot and a brief discussion about the selection of that size quadrat and what it means for inclusion of cell types in a given quadrat

``````{r WT1-rarefaction, echo=FALSE, warning=FALSE, message=FALSE, results='hide',fig.show='hide'}

# note that you may DO the analysis in one place, and save the output plot to be generated elsewhere (in results)
####Must use total cell types.  I need to remove the space outside the thymus before I can do this.
mat <-read.table("Totalcounts100_Thymus1_2.txt", head=T, sep="\t")

# an example 10x10 matrix
#mat <- matrix(1:100, 10)
nc <- ncol(mat) # number of columns
nr <- nrow(mat) # number of rows

size <- 5 # size of the subset matrix
nmat <- 8 # number of submatrices

# sample indices of submatrices
#set.seed(1)
idxc <- sample(seq(2, nc - size), size = nmat, replace = TRUE)
idxr <- sample(seq(2, nr - size), size = nmat, replace = TRUE)

# create a list of 8 submatrices
res <- mapply(function(x, y) mat[seq(x, x + size - 1), seq(y, y + size - 1)],
              idxr, idxc, SIMPLIFY = FALSE)

# calculate the average cell values
mean(unlist(res))

```


6. Now get into the methods you already have in your R script, first setting K=2 and removing all quadrats that fall outside of the thymus from further analysis. You can do this using the subset command I think. We can then talk about more ways to maintain the identity of a given cluster of spatial quadrats (identity being the type of mouse, the section of thymus, and so on)


```{r clustering, cache=TRUE, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
library (vegan)
library(labdsv)

# This step normalizes your data and is optional.
#spe.std <- decostand(spe, "normalize")  #You can also use "standardize". See 'help' for details.

# Do the k-means clustering [read 'help' for the 'kmeans' function to see what the arguments "centers" (clusters or k) and "nstart" (randomizations) mean].

spe.kmeans_Mutant1_All <- kmeans(result0, centers=cluster, nstart=10)
spe.kmeans_Mutant1_S1 <- kmeans(result1, centers=cluster, nstart=10)
spe.kmeans_Mutant1_S2 <- kmeans(result2, centers=cluster, nstart=10)
spe.kmeans_Mutant1_S3 <- kmeans(result3, centers=cluster, nstart=10)

```

```{r assigning clusters}

library(gplots) ###Call 1 set at a time; produces image of section

v1 <- spe.kmeans_Mutant1_S1$cluster  #change this to section you are interested in looking at

#load("thymus.Rdata")
#image(v2) # make pic
v2 = matrix(v1, nrow = 100, ncol=100)
v2<-v2[1:40,1:100] # trim
image(v2) # make new pic

heatmap.2( v2, Rowv=FALSE, Colv=FALSE, dendrogram='none', cellnote=v2,
           notecol="black", trace='none', key=FALSE,lwid = c(.01,0.99),lhei = c(.01,.99),
           margins = c(0,0),col=c("green", "yellow", "blue", "red")) 

clust.ext<-v2[1,1] # get cluster number of corner pixel (should always represent 'exterior')
idx<-which(v2==clust.ext,arr.ind=T) # get addresses of all pixels of that number
#mat[mat < 0.1] <- NA
idx2<-which(idx[,1]<40) # find such pixels for which there is another pixel immediately to right
idx<-idx[idx2,] # use only those pixels
idx[,1]<-idx[,1]+1 # find pixels immediately to right
z<-table(v2[idx]) # make sorted table of cluster numbers of these pixels
clust.int<-as.integer(which(z==max(z[z!=max(z)]))) # get the cluster number that is 2nd most common (this is the thing most often 'touching' exterior, i.e. interior but not medular)
#mat[mat < 0.1] <- NA

clust.med1<-as.integer(which(z==min(z[z!=min(z)])))
clust.med2<-as.integer(which(z==min(z)))
idx.med1<-which(v2==clust.med1)
idx.med2<-which(v2==clust.med2)

mean(toFill[idx.med1,1],na.rm=T) #looking at certain column numbers
mean(toFill[idx.med2,1],na.rm=T)
mean(toFill[idx.med1,22],na.rm=T)
mean(toFill[idx.med2,22],na.rm=T)
mean(toFill[idx.med1,25],na.rm=T)
mean(toFill[idx.med2,25],na.rm=T)

col1.01<-1*(toFill[,1]>0)
col22.01<-1*(toFill[,22]>0)
col25.01<-1*(toFill[,25]>0)

mean(col1.01[idx.med1],na.rm=T)
mean(col1.01[idx.med2],na.rm=T)
mean(col22.01[idx.med1],na.rm=T)
mean(col22.01[idx.med2],na.rm=T)
mean(col25.01[idx.med1],na.rm=T)
mean(col25.01[idx.med2],na.rm=T)

heatmap.2( v2, Rowv=FALSE, Colv=FALSE, dendrogram='none', cellnote=v2,
           notecol="black", trace='none', key=FALSE,lwid = c(.01,0.99),lhei = c(.01,.99),
           margins = c(0,0),col=c("green", "yellow", "blue", "red")) 

#change the colors to reflect the number of clusters.  Can also reorder so colors are consistent between sections.

```

```{r assigning clusters, cache=TRUE, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}

if(spe.kmeans_Mutant1_All$cluster="blank", exclude="blank")
  vector(spe.kmeans_Mutant1_All$cluster,exclude="blank")

factor(spe.kmeans_Mutant1_All$cluster, exclude="blank"


```

Next Step: Figuring out how clustering is done

```{r notsurewhat, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
#dimnames(spe.kmeans$centers)[2]
#unique(spe.kmeans$cluster)
```

```{r cluster subplots, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.show='hide'}
my.order<-c(1,2,3,4) # define the order we want to plot panels
par(mfrow=c(2,2)) # make 4 subplots in 2x2 style
for (i in 1:max(spe.kmeans_Mutant1_All$cluster)){ 
  barplot(colSums(result0[which(spe.kmeans_Mutant1_All1$cluster==i),]),main=i,ylim=c(0,12000)) 
  # we pick out desired cluster and plot
}

#my.order<-c(1,2,3,4) # define the order we want to plot panels
#par(mfrow=c(2,2)) # make 4 subplots in 2x2 style
#for (i in 1:max(spe.kmeans_Mutant1_S1$cluster)){ 
#  barplot(colSums(result1[which(spe.kmeans_Mutant1_S1$cluster==i),]),main=i,ylim=c(0,12000)) 
  # we pick out desired cluster and plot
#}

#my.order<-c(1,2,3,4) # define the order we want to plot panels
#par(mfrow=c(2,2)) # make 4 subplots in 2x2 style
#for (i in 1:max(spe.kmeans_Mutant1_S2$cluster)){ 
#  barplot(colSums(result2[which(spe.kmeans_Mutant1_S2$cluster==i),]),main=i,ylim=c(0,12000)) 
  # we pick out desired cluster and plot
#}

#my.order<-c(1,2,3,4) # define the order we want to plot panels
#par(mfrow=c(2,2)) # make 4 subplots in 2x2 style
#for (i in 1:max(spe.kmeans_Mutant1_S3$cluster)){ 
#  barplot(colSums(result3[which(spe.kmeans_Mutant1_S3$cluster==i),]),main=i,ylim=c(0,12000)) 
  # we pick out desired cluster and plot
#}

```

Looking at cellular relationships between clusters
We can make a running sum of all the counts of all cell types by cluster

```{r counts, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.show='hide'}
b2<-colSums(result0[which(spe.kmeans_Mutant1_All$cluster==2),])
b3<-colSums(result0[which(spe.kmeans_Mutant1_All$cluster==3),])
b4<-colSums(result0[which(spe.kmeans_Mutant1_All$cluster==4),])

#b2_S1<-colSums(result1[which(spe.kmeans_Mutant1_S1$cluster==2),])
#b3_S1<-colSums(result1[which(spe.kmeans_Mutant1_S1$cluster==3),])
#b4_S1<-colSums(result1[which(spe.kmeans_Mutant1_S1$cluster==4),])

#b2_S2<-colSums(result2[which(spe.kmeans_Mutant1_S2$cluster==2),])
#b3_S2<-colSums(result2[which(spe.kmeans_Mutant1_S2$cluster==3),])
#b4_S2<-colSums(result2[which(spe.kmeans_Mutant1_S2$cluster==4),])

#b2_S3<-colSums(result3[which(spe.kmeans_Mutant1_S3$cluster==2),])
#b3_S3<-colSums(result3[which(spe.kmeans_Mutant1_S3$cluster==3),])
#b4_S3<-colSums(result3[which(spe.kmeans_Mutant1_S3$cluster==4),])

###and normalize these so that each cluster type has the same total amount of cells (in other words, we're ###getting the proportion of cell types in each cluster)

###```{r normalize, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.show='hide'}
b2<-b2/sum(b2)
b3<-b3/sum(b3)
b4<-b4/sum(b4)

#b2_S1<-b2/sum(b2_S1)
#b3_S1<-b3/sum(b3_S1)
#b4_S1<-b4/sum(b4_S1)

#b2_S2<-b2/sum(b2_S2)
#b3_S2<-b3/sum(b3_S2)
#b4_S2<-b4/sum(b4_S2)

#b2_S3<-b2/sum(b2_S3)
#b3_S3<-b3/sum(b3_S3)
#b4_S3<-b4/sum(b4_S3)

#We can look at the difference between pairs of clusters to see which cell types are most different between #clusters


par(mfrow=c(2,2))  ##All
barplot(b2-b3,main="2-3")
barplot(b2-b4,main="2-4")
barplot(b3-b4,main="3-4")

#par(mfrow=c(2,2))
#barplot(b2-b3,main="2-3")
#barplot(b2-b4,main="2-4")
#barplot(b3-b4,main="3-4")

#par(mfrow=c(2,2))
#barplot(b2-b3,main="2-3")
#barplot(b2-b4,main="2-4")
#barplot(b3-b4,main="3-4")

#par(mfrow=c(2,2))
#barplot(b2-b3,main="2-3")
#barplot(b2-b4,main="2-4")
#barplot(b3-b4,main="3-4")

```
The way to read the last plots is for something like "2-3", positive bars mean that cell type is more common in cluster 2 (than 3); negative bars mean the opposite. The higher the absolute number on the y-axis, the greater the difference is in that cell count between the two clusters. I think (as John suggested), that we can move ahead and be more formal with this analysis (e.g. Bray Curtis dissimilarity), but this is a useful first step.

6b. bray curtis

6c. significant diffs between cell types?

## Bray Curtis

```{r braycurtis, message=FALSE, comment="", echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
library(vegan) 
library(dplyr) # used for chaining and manipulation  
# http://blog.rstudio.org/2014/01/17/introducing-dplyr/
#reads as: take only cluster 1 data from 'spe', then take column sums (i.e. total cell type count) 
#then transpose (swap rows/cols needed for bray curtis calc)

df<-result0[which(spe.kmeans_Mutant1_All$cluster==1),] %>% colSums() %>% t()   # example of chaining. 
for (j in 2:4){  
  tmp<-result0[which(spe.kmeans_Mutant1_All$cluster==j),] %>% colSums() %>% t() 
  df<-rbind(df,tmp)
}

#df1.1<-spe1.1[which(spe.kmeans1.1$cluster==1),] %>% colSums() %>% t()
#for (j in 2:4){  
#  tmp1.1<-spe1.1[which(spe.kmeans1.1$cluster==j),] %>% colSums() %>% t() 
#  df1.1<-rbind(df1.1,tmp1.1)
#}

#df1.2<-spe1.2[which(spe.kmeans1.2$cluster==1),] %>% colSums() %>% t()
#for (j in 2:4){  
#  tmp1.2<-spe1.2[which(spe.kmeans1.2$cluster==j),] %>% colSums() %>% t() 
#  df1.2<-rbind(df1.2,tmp1.2)
#}

#df1.3<-spe1.3[which(spe.kmeans1.3$cluster==1),] %>% colSums() %>% t()
#for (j in 2:4){  
#  tmp1.2<-spe1.3[which(spe.kmeans1.3$cluster==j),] %>% colSums() %>% t() 
#  df1.3<-rbind(df1.3,tmp1.3)
#}


df<-apply(df,2,as.integer) %>% as.data.frame()
#df1.1<-apply(df1.1,2,as.integer) %>% as.data.frame()
#df1.2<-apply(df1.2,2,as.integer) %>% as.data.frame()
#df1.3<-apply(df1.3,2,as.integer) %>% as.data.frame()
```

```{r dendrogram}
###This is where I import the other files? I don't think I will be able to compare the sections with ALL

#need same number of columns to bind them.
colnames(df2)<-colnames(df)
#colnames(df1.1)=colnames(df2.1)
#colnames(df1.2)<-colnames(df2.1)
#colnames(df1.2)<-colnames(df2.1)

#dfTOTAL<-rbind(df1.2,df2.2, df1.2, df1.2, df1.1, df1.1)

BrayCurtis<-vegdist(dfTOTAL,method="bray")
#print(BrayCurtis)
hc<-hclust(BrayCurtis)
plot(hc,labels=dfTOTAL$rownames)
```

7. also Andrew Sornborger's approach for separating out the geometry/geography of clusters., and/or using connected component labeling to identify size, centroid, distance of centroids among clusters

8. jackknifing : removing one cell type at a time

# Results

1. How many mice of each type, how many replicates of mouse x thymus x prep method. ###Should this be methods?
It is critical to understand how general observed patterns compare across individuals of the same age and genotype.  Data will be obtained from multiple thymii.  variation among individuals under specific test conditions. 
2. Size of quadrat selected, and what this says about number of individual quadrats examined, number of total cell counts. People love those sorts of numbers!
3. Clustering and K selection
4. Using BC to validate how different clusters are, importance of composite versus individual layers.
5. Geography: Sornborger
6. etc.

##K means clustering reproduces previously identified regions and identifies cryptic information


K values of 3 to - were selected based on previously identified information. For K=2, the cortex and medullary regions were reproduced.  Increasing K to 3 resulted in subdividing of the medullary region. What is driving the clustering?  Dendritic cells found in CMJ? Removal of cell types? 

##Bray Curtis Identifies Degree of Similarity

Bray Curtis values ranged from BLAH to BLAH.  Talk about which samples, regions, subregions were most similar, which sections, did mutants group together? etc.  

##Geography

How do the mutant and WT differ orgazationally? Can we prove they are different statistically? Want to quantify extent to which mutants are significantly less structured than wild-type

# Discussion

To an extent, as K goes up it may be that spatial weighting is important, but we choose to avoid the problems that weighting brings up; it may also be possible that some of the areas defined as K approaches 10+ are legitimate microenvironments and this will be addressed in a later paper.

1. I can generate a map of the thymus that supports previous discussion of the geography of the thymus, but it is automatic and dependent on objective criteria of sorting the functional cell types in a spatially explicit way.
2. We believe we have identified subregions in the medulla. This is cool.
3. Though our replication is minimal, we can say that there is/is not quantifiable distinction between the spatial organization of WT and Aire- mice. This has been debated in the literature and we resolve it.
4. There is much more we can do with this approach. 
5. Ecologically, it is unusual to have the problem of such complete sampling of an ecosystem. Our approach to sampling this has strengths and concerns, generally associated with K-means clustering which has its subjective/ambiguous problems but is the best we can do!


Why this approach is novel 
Coolest thing we found, how to explain it
If we did not find a big shift in organization between treatment, it is probably because what was turned off in mutant was not a 'foundational species' in this system (see http://bioscience.oxfordjournals.org/content/61/10/782.full ) as such is a minor perturbation for organization of system, even if ecosystem service ie functionality of thymus is affected. We have to separate these effects after all! It is also important to note that the cell types are not independent arrivals in the thymic ecosystem. As a developmental process, the analogy is not complete: this would be like a coral reef developing with functionally distinct polyps spatially arranged to improve net fitness (i.e. reproductive capacity); this of course does happen in certain colonial species (hydrozoans for example) and as such this may generate its own numerical/mathematical patterns distinct from the standards of community ecology (e.g. species-area hypothesis).
Second coolest thing
Concern we would like to address

Next step for application of this approach - also bigger picture stuff.
Comparison - a manipulated salt marsh ecosystem and how B-C was used to compare treatment and sites
http://www.mass.gov/eea/docs/czm/habitat/wetlands/cape-cod-report-v2.pdf
another
http://books.google.com/books?id=Kpi9TjDQO8QC&pg=PA28&lpg=PA28&dq=bray-curtis+salt+marsh&source=bl&ots=iiISZsY9Jf&sig=YGSLp0cG06oKxB1Ung56AS2EQgw&hl=en&sa=X&ei=M44yVP-_CYylyASMioLgCQ&ved=0CDgQ6AEwAw#v=onepage&q=bray-curtis%20salt%20marsh&f=false
useful comparison because of similar taxonomic diveristy (?)
http://www.scielo.br/scielo.php?pid=S1516-89132009000600013&script=sci_arttext

http://www.researchgate.net/publication/259146839_Dissimilarity_in_plant_species_diversity_between_salt_marsh_and_neighboring_environments_decreases_as_environmental_harshness_increases

http://www.tandfonline.com/doi/full/10.1080/19425120.2014.893467#.VDKQQkuoVss
there are many more. we can learn from the methods..

# Literature Cited

assume we have vector (e.g. 10000 pixels, numbered)