---
title: "Thymus Macroecology Template"
author: "Mary Browning, Nancy Manley, John Wares"
date: "November 19, 2014"
output: word_document
bibliography: thymus.bib
---

# Introduction

You already have some text that could start to go here, in the Google Doc. Note that flagging a manuscript section with a hashtag, space, and name makes it a subset of the document that can be toggled open or closed by the little arrow next to the line number [@aguinaldo97;@gordon11].

# Methods

## Terminology

I suggest an explanatory box for readers so that quadrat, section, composite, etc. can all be explained clearly in one place

1. Where the mice come from and other necessary information about the Aire mutant line
2. Information about the antibody stains, the frozen versus paraffin protocols
3. Staining protocol and visualization with Cell Profiler: you need to explain whatever settings/parameters were used to obtain your data files, including how overlapping sections were lined up.
4. Data Transformation

```{r Data Input}
# note that putting a name after the {r } NAMES the chunk, and that makes it easier to sort through your code using the drop-down at the bottom of this windowpane. You can also open/close chunks with the little arrow next to the starting line of the chunk.
```

5. Quadrat Selection and Rarefaction: you need to explain how you did this, and the next chunk can include the code you used to achieve this. That way, our results can include one plot and a brief discussion about the selection of that size quadrat and what it means for inclusion of cell types in a given quadrat

```{r sampling design, echo=FALSE}
#plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

6. Now get into the methods you already have in your R script, first setting K=2 and removing all quadrats that fall outside of the thymus from further analysis. You can do this using the subset command I think. We can then talk about more ways to maintain the identity of a given cluster of spatial quadrats (identity being the type of mouse, the section of thymus, and so on)

6b. bray curtis

6c. significant diffs between cell types?

7. also Andrew Sornborger's approach for separating out the geometry/geography of clusters.

8. jackknifing : removing one cell type at a time

# Results

1. How many mice of each type, how many replicates of mouse x thymus x prep method. 
2. Size of quadrat selected, and what this says about number of individual quadrats examined, number of total cell counts. People love those sorts of numbers!
3. Clustering and K selection
4. Using BC to validate how different clusters are, importance of composite versus individual layers.
5. Geography: Sornborger
6. etc.

# Discussion

To an extent, as K goes up it may be that spatial weighting is important, but we choose to avoid the problems that weighting brings up; it may also be possible that some of the areas defined as K approaches 10+ are legitimate microenvironments and this will be addressed in a later paper.

1. I can generate a map of the thymus that supports previous discussion of the geography of the thymus, but it is automatic and dependent on objective criteria of sorting the functional cell types in a spatially explicit way.
2. We believe we have identified subregions in the medulla. This is cool.
3. Though our replication is minimal, we can say that there is/is not quantifiable distinction between the spatial organization of WT and Aire- mice. This has been debated in the literature and we resolve it.
4. There is much more we can do with this approach. 
5. Ecologically, it is unusual to have the problem of such complete sampling of an ecosystem. Our approach to sampling this has strengths and concerns, generally associated with K-means clustering which has its subjective/ambiguous problems but is the best we can do!

# Literature Cited
